---
title: "Electricity_plots"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r}
library(tidyverse)
```

```{r}
#data
electricity_data <- "~/Desktop/data_electricity/nice_electricity/"
colnames <- c("data_id", "date", "load", "ev")

```

```{r}
electricity <- 
  list.files(path = electricity_data, full.names = TRUE) %>% 
  map_dfr(
    function(x) read_tsv(x, skip = 10, col_names = colnames) 
  )
```

```{r}
nice_data <- 
  read_csv("~/Desktop/data_electricity/nice_electricity/april_2017.csv", skip = 10, col_names = colnames) %>% 
  rbind(read_tsv("~/Desktop/data_electricity/nice_electricity/august_2017.txt", skip = 10, col_names = colnames)) %>% 
  rbind(read_tsv("~/Desktop/data_electricity/nice_electricity/december_2017.txt", skip = 10, col_names = colnames)) %>% 
  rbind(read_tsv("~/Desktop/data_electricity/nice_electricity/february_2017.txt", skip = 10, col_names = colnames))
```

```{r}
nice_data <- 
  nice_data %>% 
  mutate(
    date_day = mdy_hm(date),
    wday = wday(date_day, label = TRUE),
    day_of = day(date_day),
    hour = hour(date_day),
    minute = minute(date_day),
    ev = as.integer(ev)
  ) 

ev_use_day_data_id <- 
  nice_data %>% 
  group_by(day_of, data_id) %>% 
  summarise(sum_ev_use = sum(ev, na.rm = TRUE)) %>% 
  ungroup()

nice_data <- 
  nice_data %>% 
  left_join(ev_use_day_data_id, by = c("day_of", "data_id")) %>% 
  mutate(ev_use = if_else(sum_ev_use > 1, "yes", "no"))
```



```{r}
nice_data %>% 
  filter(load < 30) %>% 
  mutate(time = hour + minute/60) %>% 
  group_by(time, ev_use) %>% 
  summarise(median_load = median(load, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(time, median_load, color = ev_use)) +
  geom_line()

nice_data %>% 
  mutate(time = hour + minute/60) %>% 
  filter(load < 30) %>% 
  group_by(time, ev_use) %>% 
  summarise(avg_load = mean(load, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(time, avg_load, color = ev_use)) +
  geom_line() +
  labs(
    x = "Time of Day (Hour)",
    y = "Average Load (kW)"
  )
```

```{r}
nice_data %>%
  filter(data_id == 774) %>% 
  filter(load < 20) %>% 
  ggplot(aes(load)) +
  geom_histogram(bins = 100)
```



```{r}
nice_data %>% 
  mutate(time = hour + minute/60) %>% 
  filter(load < 30) %>% 
  group_by(time, ev_use, wday) %>% 
  summarise(median_load = mean(load, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(time, median_load, color = ev_use)) +
  geom_line() +
  facet_wrap(~wday) +
   labs(
    x = "Time of Day (Hour)",
    y = "Average Load (kW)"
  )
```





```{r}
nice_data %>% 
  group_by(time, ev_use) %>% 
  summarise(median_load = median(load, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(time, median_load, color = ev_house)) +
  geom_line()
```



```{r}
nice_data %>% 
  mutate(day_of = day(date_day)) %>% 
  mutate(ev = as.integer(ev)) %>% 
  ggplot(aes(ev)) +
  geom_histogram()
```

```{r}
nice_data %>% 
  mutate(time = hour + (minute/60)) %>% 
  group_by(time, ev_house) %>% 
  summarise(median_load = median(load, na.rm = TRUE)) %>% 
  ungroup() %>% 
  ggplot(aes(time, median_load, color = ev_house)) +
  geom_line()
  
```


#visualizing the latent space
```{r}
library(readxl)
#data
pecan_survey_data <- "~/Downloads/pecan_surveys.xlsx"
ev_info_data <- "~/Downloads/ev_data.csv"
households_data <- "~/Downloads/houseIds.txt"
val_latent_mu_data <- "~/Downloads/val_latent_mu.csv"
val_latent_var_data <- "~/Downloads/val_latent_var.csv"
households_val <- "~/Downloads/houseIdsVal.txt"
households_train <- "~/Downloads/houseIdsTrain.txt"
households_test <- "~/Downloads/houseIdsTest.txt"

#read data
pecan_survey <- read_excel(pecan_survey_data)
ev_info <- read_csv(ev_info_data)
households <- read_tsv(households_data, col_names = FALSE)
val_latent_mu <- read_csv(val_latent_mu_data)
val_latent_var <- read_csv(val_latent_var_data)
h_train <- read_tsv(households_train, col_names = FALSE)
h_test <- read_tsv(households_test, col_names = FALSE)
h_val <- read_tsv(households_val, col_names = FALSE)
```

```{r}
households %>% count(X1) %>% arrange(desc(n))
```

#plots of latent mu
```{r}
val_latent_mu %>% 
  ggplot(aes(`mu-01`, `mu-02`)) +
  geom_point(alpha = .4)

val_latent_mu %>% 
  ggplot(aes(`mu-02`, `mu-03`)) +
  geom_point(alpha = .4)

val_latent_mu %>% 
  ggplot(aes(`mu-01`, `mu-03`)) +
  geom_point(alpha = .4)

val_latent_mu %>% 
  ggplot(aes(`mu-03`, `mu-04`)) +
  geom_point(alpha = .4)

val_latent_mu %>% 
  ggplot(aes(`mu-04`, `mu-05`)) +
  geom_point(alpha = .4)

```

#plots of latent var
```{r}
val_latent_var %>% 
  ggplot(aes(`var-01`, `var-02`)) +
  geom_point(alpha = .4)
```

#plots of latent var with latent mu
```{r}
val_latent_var %>% 
  cbind(val_latent_mu) %>% 
  ggplot(aes(`var-01`, `mu-01`)) +
  geom_point(alpha = .4)

```

#pecan survey
```{r}
pecan_survey 
```

#ev_info
```{r}
ev_info %>% 
  count(vehicle_type) %>% 
  filter(
    !is.na(vehicle_type), 
    n > 1
  ) %>% 
  mutate(vehicle_type = fct_reorder(vehicle_type, n)) %>% 
  ggplot(aes(vehicle_type, n)) +
  geom_col() +
  labs(
    x = "Vehicle Type",
    y = "Count"
  )
```



```{r}
library(corrplot)
summaries_train <- read_csv("~/Downloads/some_results/car_vae_12/summaries_train.csv")
summaries_val <- read_csv("~/Downloads/some_results/car_vae_12/summaries_val.csv")
#val_mu <- read_csv("~/Downloads/some_results/car_vae_12/val_latent_mu.csv")
#val_var <- read_csv("~/Downloads/some_results/car_vae_12/val_latent_var.csv")
val_mu <- read_csv("~/Downloads/vae_results/ev_car/val_latent_mu.csv")
val_var <- read_csv("~/Downloads/vae_results/ev_car/val_latent_var.csv")
ev_meta <- read_csv("~/Downloads/ev_metadata.csv")
```

```{r}
full_ev1 <- 
  ev_meta %>% 
  filter(any_electric_vehicle == "Yes") %>% 
  mutate(
    vehicle_chargn_home = str_replace(vehicle_chargn_home, "%", ""),
    vehicle_chargn_home = as.numeric(vehicle_chargn_home),
    vehicle_chargn_level2ev = str_replace(vehicle_chargn_level2ev, "%", ""),
    vehicle_chargn_level2ev = as.numeric(vehicle_chargn_level2ev),
    vehicle_chargn_public = str_replace(vehicle_chargn_public, "%", ""),
    vehicle_chargn_public = as.numeric(vehicle_chargn_public),
    vehicle_chargn_work = str_replace(vehicle_chargn_work, "%", ""),
    vehicle_chargn_work = as.numeric(vehicle_chargn_work)
  ) %>% 
  left_join(
    h_val %>% 
    cbind(val_mu) %>% 
    cbind(val_var),
    by = c("dataid" = "X1")
  )

cor_m <- 
  full_ev1 %>% 
  select(vehicle_chargn_home : vehicle_chargn_work, `mu-01` : `var-05`) %>% 
  cor(method = "pearson", use = "complete.obs")

#corrplot(cor_m, method = "circle", type="upper", order="hclust", tl.srt=45)
```


```{r}
library(viridis)
cor_m %>% 
  as.tibble() %>% 
  mutate(correlated_variable = colnames(cor_m)) %>% 
  select(correlated_variable, `mu-01` : `var-05`) %>% 
  filter(correlated_variable %in% c("vehicle_chargn_home", "vehicle_chargn_level2ev", "vehicle_chargn_public", "vehicle_chargn_work")) %>% 
  gather(-correlated_variable, key = "latent_variable_stat", value = "correlation") %>% 
  ggplot(aes(correlated_variable, latent_variable_stat, fill = correlation)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    x = "Electric Vehicle Variable",
    y = "Latent Variable Statistic",
    title = "Correlation of our VAE Latent Space \n with Electric Vehicle characteristics"
  ) +
  scale_fill_viridis() 
```


```{r}
full_ev1 %>% 
  ggplot(aes(`mu-01`, `mu-02`, color = vehicle_chargn_level2ev)) +
  geom_point() +
  scale_color_viridis() 
  #labs(  title = "EV Latent Spaces and the Percent \nof Level 2 EV Charging")

full_ev1 %>% 
  ggplot(aes(`mu-01`, `mu-02`, color = vehicle_chargn_home)) +
  geom_point() +
  scale_color_viridis() 
  #labs(title = "EV Latent Spaces and the Percent \nof EV Charging that happens at Home")
```


```{r}
cor_m %>% 
  as.tibble() %>% 
  mutate(correlated_variable = colnames(cor_m)) %>% 
  select(correlated_variable, `mu-01` : `var-05`) %>% 
  filter(correlated_variable %in% c("vehicle_chargn_home", "vehicle_chargn_level2ev", "vehicle_chargn_public", "vehicle_chargn_work")) %>% 
  gather(-correlated_variable, key = "latent_variable_stat", value = "correlation") %>% 
  arrange(desc(abs(correlation)))
```

#Calculate the mean of the latent variables and then see how they relate to our covariates
```{r} 
mean_latent <- 
  full_ev1 %>% 
  group_by(dataid) %>% 
  summarise(
    mean_m1 = mean(`mu-01`),
    mean_m2 = mean(`mu-02`),
    mean_m3 = mean(`mu-03`),
    mean_m4 = mean(`mu-04`),
    mean_m5 = mean(`mu-05`),
    vehicle_chargn_home = mean(vehicle_chargn_home),
    vehicle_chargn_level2ev = mean(vehicle_chargn_level2ev),
    vehicle_chargn_public = mean(vehicle_chargn_public),
    vehicle_chargn_work = mean(vehicle_chargn_work)
  ) %>% 
  ungroup() %>% 
  filter(!is.na(mean_m1)) %>% 
  replace(., is.na(.), 0)

cor_mean <- 
  mean_latent %>% 
  select(-dataid) %>% 
  cor(method = "pearson", use = "complete.obs")

corrplot(cor_mean, method = "circle", type="upper", order="hclust", tl.srt=45)
```


#let's look at top 25 and bottom 25 for households and do this same approach
```{r}
latent_25_75 <- 
  full_ev1 %>% 
  select(dataid, vehicle_chargn_home:vehicle_chargn_work, `mu-01`:`var-05`) %>% 
  group_by(dataid) %>% 
  summarise(
    quant_25_m1 = quantile(`mu-01`, .25, na.rm = TRUE),
    quant_75_m1 = quantile(`mu-01`, .75, na.rm = TRUE),
    quant_25_m2 = quantile(`mu-02`, .25, na.rm = TRUE),
    quant_75_m2 = quantile(`mu-02`, .75, na.rm = TRUE),
    quant_25_m3 = quantile(`mu-03`, .25, na.rm = TRUE),
    quant_75_m3 = quantile(`mu-03`, .75, na.rm = TRUE),
    quant_25_m4 = quantile(`mu-04`, .25, na.rm = TRUE),
    quant_75_m4 = quantile(`mu-04`, .75, na.rm = TRUE),
    quant_25_m5 = quantile(`mu-05`, .25, na.rm = TRUE),
    quant_75_m5 = quantile(`mu-05`, .75, na.rm = TRUE),
    vehicle_chargn_home = mean(vehicle_chargn_home),
    vehicle_chargn_level2ev = mean(vehicle_chargn_level2ev),
    vehicle_chargn_public = mean(vehicle_chargn_public),
    vehicle_chargn_work = mean(vehicle_chargn_work)
  ) %>% 
  ungroup() %>% 
  filter(!is.na(quant_25_m1)) %>% 
  replace(., is.na(.), 0)

cor_latent_25_75 <- 
  latent_25_75 %>% 
  select(-dataid) %>% 
  cor(method = "pearson", use = "complete.obs")

corrplot(cor_latent_25_75, method = "circle", type="upper", order="hclust", tl.srt=45)
```


#Now let us do a case study of an individual and their distributions

```{r}
full_ev1 %>% 
  count(dataid) %>% 
  arrange(desc(n)) %>% 
  left_join(
    full_ev1 %>% 
    distinct(dataid, vehicle_chargn_home, vehicle_chargn_work)
  ) %>% 
  arrange(desc(n, vehicle_chargn_home))
```

```{r}
#household with 100% charging at home
full_ev1 %>% 
  filter(dataid == 4336) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value)) +
  geom_density() +
  facet_wrap(~latent_mu, nrow = 5)


full_ev1 %>% 
  filter(dataid == 4336) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value, fill = latent_mu, color = latent_mu)) +
  geom_density(alpha = 0.1) +
  coord_cartesian(xlim = c(-5, 5))
  #labs(title = "Latent distributions for household that charge 100% at home")

#household with 0% charging at home
full_ev1 %>% 
  filter(dataid == 6691) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value, fill = latent_mu, color = latent_mu)) +
  geom_density(alpha = 0.1) +
  coord_cartesian(xlim = c(-5, 5))
  #labs(title = "Latent distributions for household that charges 0% at home")

#household with 90% charging at work
full_ev1 %>% 
  filter(dataid == 26) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value, fill = latent_mu, color = latent_mu)) +
  geom_density(alpha = 0.1) +
  labs(title = "Latent variable distributions for household that charges 90% at work")
```


#Now let us look at our household latent space and household covariates

```{r}
audit_pecan <- read_csv("~/Documents/audit_pecan.csv")
#val_mu <- read_csv("~/Downloads/val_latent_mu (1).csv")
#val_var <- read_csv("~/Downloads/val_latent_var (1).csv")

val_mu <- read_csv("~/Downloads/vae_results/household/val_latent_mu.csv")
val_var <- read_csv("~/Downloads/vae_results/household/val_latent_var.csv")

house_summary_latent <- 
  h_val %>% 
  cbind(val_mu ) %>% 
  left_join(
    audit_pecan %>% 
    group_by(dataid) %>% 
    summarise(
      number_of_levels = mean(number_of_levels, na.rm = TRUE),
      no_bedrooms = mean(no_bedrooms, na.rm = TRUE),
      conditions_square_foot = mean(conditions_square_foot, na.rm = TRUE),
      house_volume_cubic_feet = mean(house_volume_cubic_feet, na.rm = TRUE)
    ) %>% 
    ungroup(),
    by = c("X1" = "dataid")
  ) %>% 
  filter(!is.na(house_volume_cubic_feet)) 


```


```{r}
house_summary_latent %>%
  select(-X1) %>% 
  cor(method = "pearson", use = "complete.obs") %>% 
  corrplot(method = "circle", type="upper", tl.srt=45)

#results show a very strong negative correlation between home size, number of rooms, and our second latent variable mu_02. mu_03 also is negatively associated as well as mu_04.
```

```{r}
house_cor <- 
  house_summary_latent %>%
  group_by(X1) %>% 
  summarise(
    mean_m1 = mean(`mu-01`),
    mean_m2 = mean(`mu-02`),
    mean_m3 = mean(`mu-03`),
    mean_m4 = mean(`mu-04`),
    mean_m5 = mean(`mu-05`),
    no_bedrooms = mean(no_bedrooms),
    number_of_levels = mean(number_of_levels),
    conditions_square_foot = mean(conditions_square_foot),
    house_volume_cubic_feet = mean(house_volume_cubic_feet)
  ) %>% 
  ungroup() %>% 
  filter(!is.na(mean_m1)) %>% 
  replace(., is.na(.), 0) 
  
house_cor %>% 
  select(-X1) %>% 
  cor(method = "pearson", use = "complete.obs") %>% 
  as_tibble() %>% 
  mutate(correlated_variable = colnames(house_cor %>% select(-X1))) %>% 
  select(correlated_variable, `mean_m1` : `mean_m5`) %>% 
  filter(correlated_variable %in% c("no_bedrooms", "number_of_levels", "conditions_square_foot", "house_volume_cubic_feet")) %>% 
  gather(-correlated_variable, key = "latent_variable_stat", value = "correlation") %>% 
  ggplot(aes(correlated_variable, latent_variable_stat, fill = correlation)) +
  geom_tile() +
  theme(axis.text.x = element_text(angle = 25, hjust = 1)) +
  labs(
    x = "Household Variable",
    y = "Latent Variable Statistic",
    title = "Correlation of the Household VAE Latent Space Means \n with House Characteristics"
  ) +
  scale_fill_viridis() 


```



#let's look at a big house and a little house

```{r}
house_summary_latent %>%
  distinct(X1, conditions_square_foot) %>% 
  arrange(desc(conditions_square_foot))
```

```{r}
#big house with square footage of 4447
house_summary_latent %>%
  filter(X1 == 9935) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value, fill = latent_mu, color = latent_mu)) +
  geom_density(alpha = 0.1) +
  coord_cartesian(
    xlim = c(-5, 5)
  )

#small house with square footage of 1080
house_summary_latent %>%
  filter(X1 == 9609) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value, fill = latent_mu, color = latent_mu)) +
  geom_density(alpha = 0.1) +
  coord_cartesian(
    xlim = c(-5, 5)
  ) 


#alternative display
house_summary_latent %>%
  filter(X1 == 9935) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value)) +
  geom_density(alpha = 0.1) +
  labs(title = "Latent variable mu distributions for largest household (4447 sqft)") +
  coord_cartesian(
    xlim = c(-5, 5)
  ) +
  facet_wrap(~latent_mu, ncol = 5)

#small house with square footage of 1080
house_summary_latent %>%
  filter(X1 == 9609) %>% 
  select(`mu-01`:`mu-05`) %>% 
  gather(key = "latent_mu", value = "value") %>% 
  ggplot(aes(value)) +
  geom_density(alpha = 0.1) +
  labs(title = "Latent variable mu distributions for smallest household (1080 sqft)") +
  coord_cartesian(
    xlim = c(-5, 5)
  ) +
  facet_wrap(~latent_mu, ncol = 5)

```





####training and validation loss plots
#plots for the training and validation
```{r}
#EV training
summaries_train <- read_csv("~/Downloads/vae_results/ev_car/summaries_train.csv")
summaries_val <- read_csv("~/Downloads/vae_results/ev_car/summaries_val.csv")

summaries_train %>% 
  mutate(Run = "Train") %>% 
  rbind(
    summaries_val %>% 
    mutate(Run = "Validation") %>% 
    slice(1:4)
  ) %>% 
  ggplot(aes(epoch, loss, color = Run)) +
  geom_line() +
  labs(
    x = "Epoch",
    y = "Loss",
    title = "EV Load VAE Training and Validation Loss over Time"
  )

#house training
summaries_val2 <- read_csv("~/Downloads/vae_results/household/summaries_val.csv")
summaries_train2 <- read_csv("~/Downloads/vae_results/household/summaries_val.csv")

summaries_train2 %>% 
  mutate(Run = "Train") %>% 
  rbind(
    summaries_val2 %>% 
    mutate(Run = "Validation") %>% 
    slice(1:4)
  ) %>% 
  ggplot(aes(epoch, loss, color = Run)) +
  geom_line() +
  labs(
    x = "Epoch",
    y = "Loss",
    title = "Overall Household Load VAE Training and Validation Loss over Time"
  )
```




